---
title: "`r knitr::current_input()`"
date: "`r paste('last updated', 
    format(lubridate::now(), ' %d %B %Y'))`"
output:   
  html_document:  
        code_folding: hide
        toc: true 
        toc_depth: 2
        toc_float: true
  pdf_document:
    latex_engine: xelatex
    toc:  TRUE
    toc_depth:  3
    extra_dependencies: ["bbm", "threeparttable"]  
fontsize: 10pt
geometry: margin=0.5in,top=0.25in
editor_options: 
  chunk_output_type: console
TAGS:  fecal fat,
---


```{r setup, include=FALSE		}
knitr::opts_chunk$set(
											echo = TRUE,  
											comment = "      ##",  
											error = TRUE, 
											collapse = TRUE)
library(data.table)
library(lubridate)   # ymd requires
```

####  Fecal_fatty_acids (g/24 hours)
  *  72 hour stool tests
  *  normal <7g/24 hours
  *  10/31/18 test was 72 hours, never corrected properly.
  *  With this correction:  the first test reported fat 34 g/day and total
    fecal weight (3 days) 834 g.
  * This works out to 102 g fat (34 g/day * 3 days) out of 834 g, or 12.4% 
  *  Next tests are done same way.
  *  Result:   I see very little variation in percent fecal fat, WITH or
    WITHOUT Zenpep.

```{r fecal_fats}

fecal  <- data.table(date=c("10/31/18", "4/2/19","5/24/21" ),
                          test_length = c(72,72,72),
                          fatty_acids = c(34.4, 44, 20),
                          total_fecal_wt = c(834, 1114, 502),
                          comments = c("72 hours don't think ever got fixed",
                                       NA, NA))
fecal


## convert date from character to numeric
dt  <- fecal[, .( date = lubridate::mdy(date), 
                 percent_fat = 100*(3*fatty_acids)/total_fecal_wt,
          test_length, fatty_acids, total_fecal_wt, comments)]

fecal  <- dt
plot(x = fecal$date, y  = fecal$percent_fat ,
     ylim=c(5.0, 20),
     main = "Percent Fecal Fat for Each 72-hour Test",
     col="black", 
     type= "p")
```
```{r end}
knitr::knit_exit()
```


#### Review Ur Oxalate (is this old?)

```{r read_data,  include=FALSE}
## read saved data, ( downloaded, cleaned from google spreadsheet)
result <- readr::read_csv(here("data_clean", "ALL_lab_data.csv"))
```

```{r check_data, include=FALSE}
## before plots, check the data
# summarize,  Count for each Test (in alpha order)
{
result %>% 
  group_by(Test_Name) %>% 
  summarize(count = n()) %>% 
  arrange(Test_Name) %>% 
  print(n=50)
}

```
##  Urinary Oxalate | Summary | Jim Rothstein (DOB: 12/23/55)

  *  Been using Zenprep for 4 weeks. Diet also changed.   24-hour urine to be
    sent to Litholink tomorrow.   Dr. Lieske suggested phone appointment 2
    weeks after I send sample, so last week in April.


### Past Data for:   Urine Oxalate 24 hour  

  *  Mix of Litholink and Kaiser NW  labs.    
  *  The value for 2/25/2014 (668) is an error and will be removed.
```{r table_urine_oxalate, echo=FALSE	}
{
ur_ox_tibble  <- 
  filter(result, Test_Name == "Ur_Ox24" )
}
ur_ox_tibble
```

```{r, remove_bad_data, include=FALSE}

# filter uses single &  ;   not &&. 
ur_ox_tibble  <- filter(result, Test_Name == "Ur_Ox24" & Test_Result < 500)

```


### UrOx vs time  
  *  Normal is < 35 mg (red line)
  *  Beginning mid-2014, Ca supplements increased to 1000+ mg per day, esp 2018
    on.

```{r plot_urOx, echo=FALSE		}

{
g <-filter(result, Test_Name == "Ur_Ox24" & Test_Result < 500) %>% 
    ggplot(aes(x=Date, y=Test_Result)) + geom_point() +
    		labs (title = "Ur_Ox24", subtitle = "2008 - 2021") 
g  <- g + geom_hline(aes(yintercept=35),  color="red")

}
g 
```
 
```{r, save_plot, include=FALSE}
# save plot

name  <- paste0("006_UrOx_report.",
                format(Sys.Date(), "%Y_%m_%d"),
                ".png")
# save last plot
ggsave( here("output", name), device=png())

```

```{r eval=FALSE	, include=FALSE	}
# purpose?

## save graph, retrieve, AS object - without recalculating

saveRDS(g,file=here("data_clean/UrOx.rds"))
g <- readRDS(here("data_clean/UrOx.rds"))
g
```

### Histogram:  Distribution of 24-hour Urinary Oxalate.
  *  Suppose all observations were done under same conditions (a big leap)
  *  Definitely not normally distributed.

```{r descriptive_stats, echo=FALSE}
m  <- mean(ur_ox_tibble$Test_Result)
sd  <- sd(ur_ox_tibble$Test_Result)
cat("mean = ", m, "\n")
cat("sd = ", sd, "\n")

hist(ur_ox_tibble$Test_Result, main = "UrOx",  xlab="UrOx mg/24 hours")
```

```{r plot_base, eval=FALSE, include = FALSE}
### Boxplot (using base)

ur_ox_tibble
ur_ox_tibble$enzyme = "pre"
ur_ox_tibble
boxplot(Test_Result ~ pre_enzyme, data=ur_ox_tibble, frame=FALSE)

# TODO - fix
x  <- cbind("pre_enzyme", ur_ox_tibble$Test_Result)
points(x)

```


### Boxplot
```{r boxplot}

p  <- ggplot(ur_ox_tibble, aes(x = "Without Enzymes", y=Test_Result))
p + geom_boxplot(varwidth= FALSE) + geom_jitter(width=0.01)

```

###  How I understand it.
  *   Oxalate has no role in human health; it is not a nutrient. Once
    consumed it must be eliminated[^1] The body's preference is via stool, ie
    before oxalate enters the blood.   Kidneys must remove the remainder.

[^1]:  I am ignoring liver production, deposits in tissue and no oxalate
    degradation (bacteria).  I have many questions!

  *  Therefore:

$$ Oxalate IN     = Oxalate OUT $$

  *  Specifically, 
  
$$Ox_{diet} = Ox_{urine} + Ox_{stool}$$

  *   One solution seems clear:   consume NO oxalate!   Unfortunately, many of
    the healthiest plants are also high in oxalate. [^2] [^3]

[^2]:   Plants  use oxalate to bind excess calcium.

[^3]: See, for example,  https://kidneystones.uchicago.edu/how-to-eat-a-low-oxalate-diet/

  *  In healthy humans oxalate eliminated in stool by combining with calcium in
    the digestive track before entering blood.  In my case, calcium
    supplementation did not work.  

  *   For me, the chief suspect is lack of pancreatic enzymes (plus excess
    dietary oxalate).  Too much oxalate is reaching blood.


### Trial using Pancrelipase (Zenprep), Lipase=25,000 USP (~ 11 per day)

  *  Began 16 MARCH 2021; also changed diet.  First Litholink expected
    mid-April.
  *  Unlike many patients, I exhibit no obvious signs of fat malabsorption.[^4]
    Therefore the benefit, if any, will be in reduced UrOx, separate from diet.

[^4]: No malnutrition, no diarrhea 

### Bottom Line:  How is "success" defined?   

  *  Specifically, how will UrOx distribution change; 
what parameters indicate "success" ?
  *  How many data points are needed to reach conclusion?



```{r knit_exit}
knitr::knit_exit()
```

render
```{r render, eval=FALSE	}

file <- "rmd/010_fecal_fats.Rmd"
file  <- basename(file)
file

rmarkdown::render(file,
                  output_format = "pdf_document",
                  #output_format = "html_document",
                  output_file = "~/Downloads/print_and_delete/out")
```
